---
title: "Basic summaries"
author: "Francisco Guzmán"
date: "May 12, 2016"
output:
  pdf_document:
    highlight: tango
    number_sections: yes
  html_document: default
---
`r library(knitr);opts_chunk$set(comment=">>>")`

In this recipe, you'll learn:
- How to look at data summaries (average, standard dev, median)
- How to create a basic histogram
- How to create a basic annotation line

# Loading data

```{r}
# Let's load some pre-defined R data
data(mtcars)
help (mtcars)
# Let's peek into the data
head(mtcars)
# Let's see its structure
str(mtcars)
```

# Let's calculate summaries of variables!
```{r}

# Let's calculate the average MPG consumption
mean(mtcars$mpg)
# Let's calculate the standard deviation
sd(mtcars$mpg)
# Shortcut, calculate  summary statistics
summary(mtcars$mpg)

# let's generate an histogram of the mpg variable with 10 bars
hist(mtcars$mpg,breaks=10,main="Histogram of MPG", xlab="Miles per Gallon (MPG)")
# Let's draw a vertical line at the mean
abline(v=mean(mtcars$mpg),col="red",lty=2)

# Calculate some bi-variate stats
#let's plot MPG vs weight
plot(mtcars$wt,mtcars$mpg, main="Automobile MPG vs Weight", ylab="Miles per Gallon (MPG)",xlab="Weight (lb/1000) ")

#There is a clear (inverse) relationship between both
#Let's calculate correlation
cor(mtcars$mpg,mtcars$wt)
```

#Exercise: 

  1- Calculate summaries for `HP` and `wt`

  2- Plot histogram of `HP`

  3- Plot scatterplot of `HP` vs `MPG` and calculate correlation



