---
title: "Summaries by factors"
author: "Francisco Guzmán"
date: "May 12, 2016"
output:
  pdf_document:
    highlight: tango
    number_sections: yes
  html_document: default
---

Let's create summaries by factors.
Here you'll learn how to:

  - Install a new package
  - Load a package
  - Get average summary statistics by group
  - Dump a latex table


# Install a package
There are two ways to install a package:
 1. Menu
 2. Command Line
 
```{r}
#install.packages('plyr')
#install.packages('xtable')
```

# Load plyr

Plyr is a library that offers the command `ddply`, a very useful tool to explore your data

```{r}
library("plyr")
```

# Let's summarize the mtcars data to learn about gas consumption


```{r}
data(mtcars)
#summarize MPG data in mtcars by number of cylinders
ddply(mtcars,.(cyl),summarize,avg_MPG=mean(mpg))

#summarize MPG data in mtcars by transmission
ddply(mtcars,.(am),summarize,avg_MPG=mean(mpg))
# Let's modify am into a factor with labels 0="AUTOMATIC",1="MANUAL"
mtcars$f_am<-factor(mtcars$am,levels=c(0,1),labels=c("AUTOMATIC","MANUAL"))
# Now, let's summarize it again by the new factor
ddply(mtcars,.(f_am),summarize,avg_MPG=mean(mpg))

#summarize MPG and weight data in mtcars by number of cylinders and transmission
my_table<-ddply(mtcars,.(cyl,f_am),summarize,avg_MPG=mean(mpg),avg_WT=mean(wt))
print(my_table)

```


```{r,results='asis'}
library(xtable)

xtab<-xtable(my_table,caption="Average Consumption per number of cylinders", label="lab:avgmpg")
print(xtab,type="latex")
```

# Exercise:
1- Change cylinders into a factor (4="SMALL",6="MEDIUM",8="BIG"), repeat the summarization and print the latex table

2- Summarize average MPG by number of gears



