---
title: "ggplot with factors"
author: "Francisco Guzman"
date: "May 12, 2016"
output:
  pdf_document:
    highlight: tango
    number_sections: yes
  html_document: default
---

In this recipe, we will learn:

  - How to load data from a text file (reinforce)
  - How to create summaries with ddply (reinforce)
  - How to plot bar plots with ggplot
  - How to plot different views using facets
  - How to save your graph as a pdf (reinforce)

## Load data
```{r load data}

df <-read.table("data/mte_metrics.dat",header=FALSE,col.names=c("metric","lang_pair","testset","system","score") )
# The data looks a bit different than in the previous exercise
head(df)
```


# Bar plots with ggplot
```{r bar plot}
# load ggplot2
library(ggplot2)

# If we just use the geom bar over the original data, it will create a frequency distribution (histogram) by the variable of interest (lang pair)
ggplot(df,aes(x=lang_pair))+geom_bar()
# But this is not very useful because we want a comparison of scores between BLEU and DR_LEX

#Let's add scores into the picture
ggplot(df,aes(x=lang_pair,y=score))+geom_bar(stat="identity") # we need to tell ggplot that we want to use the value of y "as_is" to determine the height of the bars.

#Still this picture is not useful, because we can't compare the two metrics
#Let's color the bars
ggplot(df,aes(x=lang_pair,y=score,fill=metric))+geom_bar(stat="identity")
#This is not a good graph because it is "stacking the graphs". Let's put them side by side

ggplot(df,aes(x=lang_pair,y=score,fill=metric))+geom_bar(stat="identity", position="dodge")
# This is more useful!

```
# Method 2: using ddply
We first prepare a useful summary
```{r summary}
#we use ddply to obtain average scores by metric and language pair
library(plyr)
summary<-ddply(df,.(lang_pair,metric),summarize,avg_score=mean(score))
print(summary)
```

```{r alternative plot over summaries}

ggplot(summary,aes(x=lang_pair,y=avg_score,fill=metric ))+geom_bar(stat="identity",position="dodge")

```

# Using facets

Now, instead of "dodging" the bars, we could plot them side by side using facets

```{r facetting the metric}
ggplot(summary,aes(x=lang_pair,y=avg_score,fill=metric))+geom_bar(stat="identity")+
  facet_grid(.~metric)  #this wil facet horizontally using the factor  metric
```

```{r facetting the lang pair}
ggplot(summary,aes(x=metric,y=avg_score,fill=metric))+geom_bar(stat="identity")+
  facet_grid(.~lang_pair) 
```

# Set titles, and print

```{r prepare and print}

my_plot<- ggplot(summary,aes(x=metric,y=avg_score,fill=metric))+geom_bar(stat="identity",width=0.9)+  facet_grid(.~ lang_pair) + ggtitle("Comparison of average system scores of two MTE metrics") + xlab("MTE metrics")+ ylab("Average scores over systems") +scale_fill_manual(values = c("BLEU"="blue","DR_LEX"="green"))

pdf("img/my_second_plot.pdf",height=5,width=7)
print(my_plot)
dev.off()
```

